\section{Problem Statement}
\label{sec:ProblemStatement}

\subsection{Background}

As of 2023, the top four most popular programming languages: \texttt{Python},
\texttt{JavaScript},
\texttt{TypeScript}, and \texttt{Java}, permit to various degrees a loose or optional type
system\cite{STACK}. This flexibility often results in ambiguities that may hinder the
readability and maintainability of code as type deductions occur either at compile
time or runtime. With increasing complexities in software systems, there is an
emerging need for a language and compiler that enforce discipline and explicitness in
code while offering robust control over system memory.

\subsection{Problem}

Developers currently seem to prefer dynamic or weak typing in popular programming
languages\cite{STACK} which may compromise code readability and maintainability due to implicit
type inferences. This raises challenges in discerning the intentions of the developer
solely based on the source code. Futhermore, these languages often utilize garbage
collectors, sacrificing control over system memory, and adding runtime overhead which
could have been handled at compile time.

\subsection{Objectives}
\label{sec:Objectives}

To design and develop a compiled programming language which encourages clarity, offers
type safety and memory safety while avoiding the issues typically associated with
garbage collectors. It will make use of the same technique as the \texttt{Rust} programming
language which uses a borrow checking and reference based system to ensure memory
safety\cite{RUST}. To ensure type safety, our language won't make use of type
inference, meaning the compiler will never attempt to guess what type a developer
intends to use. If the \typeChecker{} cannot identify the type of an expression at
compile time it will instead throw an error. \\

The following objectives can be summized:

\begin{itemize}
  \item Development of a compiled programming language.
  \item Explicitness in code.
  \item Type safety at compile time.
  \item Memory safety without garbage collection.
\end{itemize}
